---
title: "Assignment11"
author: "James B"
date: "17/11/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(palmerpenguins)
library(tibble)
```

# Part 1
Explain the following concepts:

1. Statistical hypothesis: Frames the research question we want to answer in terms of the parameters of a statistical model. It enables us to answer the research question by testing it on the basis of observed data modeled as the realised values taken by a collection of random variables.

2. Null hypothesis: The default position we take when attempting to carry out a hypothesis test, typically we declare the absence of any phenomena. 

3. Alternative hypothesis: Something of interest that differs from the null, generally a phenomena we wish to prove 

4. Test statistic: Is a function of the data which has two characteristics:
a) Has a known distribution under the null hypothesis $H_0$, e.g. a gaussian distribution.
b) Often takes on large or extreme values under the alternative hypothesis $H_1$.
The test statistic takes on typical values for $H_0$ -> we accept the null hypothesis $H_0$.
The test statistic takes on non-typical values for $H_1$ -> we reject the null & accept the alternative $H_1$. 

5. Type 1 error: Where we reject the null hypothesis despite it being actually correct. We are generally more cautious about Type 1 errors than Type 2 errors since the null hypothesis is our default.

6. Type 2 error: Where we fail to reject the null hypothesis when the alternative is actually correct.

7. The size of a test: The test size of a test $\alpha_{test}$ is the probability of Type 1 error under the null hypothesis:
$\alpha_{test} = P(Type\ 1\ error\ |\ H_0\ is\ true)$

8. The power of a test: The power of a test is 1 - $\beta_{test}$ where $\beta_{test}$ is the probability of Type 2 error under the alternative: $\beta_{test} = P(Type\ 2\ error\ |\ H_1\ is\ true).$

9, The significance level: The significance of a test $\alpha$ is an upper bound on the test size $\alpha_{test} \leq \alpha$. A valid hypothesis test requires that the significance level be chosen in advance of seeing the data. The lower the significance $\alpha$, the lower the power 1 - $\beta_{test}$. It is the probability of rejecting the null hypothesis give that it is true (a type 1 error), usually set at 5%.

10. The p-value: The probability of obtaining test results at least as extreme as the results actually observed, under the assumption that the null hypothesis $H_0$ is correct. Given test statistic $\hat T$, it is the probability that $\hat T$ is at least as extreme as $\tau$ under the null $H_0$.

# Part 2
```{r part2, echo=FALSE}
bill_adelie <- penguins %>%
  filter(species=="Adelie") %>%
  pull(bill_length_mm) # extract the species and the column we want

tibble(bill_adelie) %>% ggplot(aes(x=bill_adelie))+geom_density()+
  theme_bw()+labs(x="Weight (grams)", y="Density") # plot density distribution

tibble(bill_adelie) %>% ggplot(aes(sample=bill_adelie))+stat_qq()+
  stat_qq_line(color="blue")+theme_bw() # plot qq plot

t.test(x=bill_adelie, mu=40, conf.level=0.99) # perform hypothesis test with mean of 40 and 99% confidence
```
For this hypothesis test we must assume that the data is identically and independently distributed, typically in a gaussian pattern, by observing the above graphs we can indeed confirm that these assumptions are well met. 

The hypothesis test carried out rejects the null hypothesis that the average bill length is 40mm, when it is actually significantly lower.

# Part 3
```{r part3, echo=FALSE}
library(Stat2Data)
data("Airlines") # load data library

delta_arrivals <- Airlines %>%
  filter(airline == "Delta", airport == "ORD") %>% # filter out specifically the rows we want
  pull(OnTime) # extract column

delta_arrivals <- as.vector(delta_arrivals) #vectorize

sample_size <- length(delta_arrivals) # sample size
num_successes <- length(which(delta_arrivals == "yes")) # count number of yes results

binom.test(x=num_successes, n=sample_size, p=0.875, alternative="two.sided") # carry out binomial test with p of 0.875
``` 
The calculated p value is given as 0.0776 which is greater than the significance level of 0.05 so it can be concluded that we accept the null hypothesis $H_0$ and reject the alternative hypothesis $H_1$. The value of 0.875 is also within the 95% confidence interval as well confirm our results.

The assumptions for a binomial hypothesis test are the following:

1) Items are dichotomous, i.e. for each item there are two outcomes only, and the data type is nominal / categorical, in this case it refers to whether a plane was on time or not.

2) The sample size is significantly less than the population

3) The sample is a fair representation of the population

4) Sample items are independent and identically distributed

# Part 4
```{r part4, echo=FALSE}
# a function that carries out a two sided hypothesis test taking in data and a null hypothesis value
two_sided_t_test <- function(data, hypo_val){  
  sample_size <- length(data) # sample size
  sample_mean <- mean(data, na.rm=TRUE) # sample mean
  sample_sd <- sd(data, na.rm=TRUE) # sample sd
  test_stat <- (sample_mean-hypo_val)/(sample_sd/sqrt(sample_size)) # calculate the test statistic
  return (2*(1-pt(abs(test_stat), df=sample_size-1))) # return the p value
}

bill_adelie <- penguins %>%
  filter(species=="Adelie") %>%
  pull(bill_length_mm) # extract the species and the column we want

two_sided_t_test(bill_adelie, 40) # use function to get output
t.test(x=bill_adelie, mu=40)$p.value

```
Here we can see that the two values are nearly identical, showing that the function has been implemented corrected